---
title: "Clusterings"
author: "Anneke Speijers"
date: "25 June 2016"
output: html_document
---


```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height= 8}
# producing differenc clusterings using different variable subsets

library(data.table)
library(ggplot2)
library(grid)
library(gridExtra)

setwd("~/Desktop/BGSE/Term3/MasterProject/GSE")

# read in data
stores_dummies_mktsharesales <- readRDS("stores_dummies_mktsharesales.RData")
stores_dummies_mktshareshelf <- readRDS("stores_dummies_mktshareshelf.RData")

# remove store dummy and total sell-in columns from one table (since they are common to both)
stores_dummies_mktshareshelf <- stores_dummies_mktshareshelf[,1:244, with=FALSE]
stores_dummies_mktshareshelf[,store_total_quantity:=NULL]

# rename shelf space columns
names(stores_dummies_mktshareshelf)[3:ncol(stores_dummies_mktshareshelf)] <- paste0(names(stores_dummies_mktshareshelf)[3:ncol(stores_dummies_mktshareshelf)], "_ss")

# combine data into one table
stores_dummies_mktshares <- merge(stores_dummies_mktsharesales, stores_dummies_mktshareshelf, by="storeID") 
rm(stores_dummies_mktsharesales, stores_dummies_mktshareshelf)

# write function to produce clusters and plot
# plot_title is a string stating which variables are included in the clustering and which are not
# features is a dataframe where each column is a store variable.
clustering <- function( plot_title, features, max_clusters=20, max_iters=100 ) {
    
    # Calculate the within group sum of squares for different k
    wss <- (nrow(features)-1)*sum(apply(as.data.frame(features),2,var))
    
    for (i in 2:20) {
        wss[i] <- sum( kmeans(features, centers=i, iter.max=max_iters)$withinss )
    }
    
    df <- data.frame(Clusters = 1:20, wwss=wss)
    clust_plot <- ggplot(data=df, aes(x=Clusters, y=wwss, group=1)) + 
        geom_line(colour="dodgerblue2") +
        geom_point(colour="dodgerblue2", size=1 ) +
        #geom_vline(xintercept = 7, linetype=4) +
        theme_bw() +
        theme(plot.title = element_text(size = rel(0.5)),
              axis.title.y = element_text(size = rel(0.5)),
              axis.title.x = element_text(size = rel(0.5)),
              axis.text.x  = element_text(size= rel(0.4)),
              axis.text.y = element_text(size= rel(0.4)) ) +
        ylab("Within groups sum of squares") +
        xlab("Number of clusters") +
        ggtitle( paste0("Clustering variables: \n", plot_title) ) 
    
    return(list(clusters=df, plot=clust_plot))
 
}

# save storeID's then remove from dataset
storeIDs <- as.integer(as.character(stores_dummies_mktshares$storeID))
stores_dummies_mktshares[,storeID:=NULL]

################################################################################
   # Clusterings using all mkt share variables (ie prod, subFam, fam,...etc) #
################################################################################

# all mkt share by sell-in variables
title1  <- "mkt share sell-in (no total sell-in or store dummies)"
feat1   <- stores_dummies_mktshares[,1:242, with=FALSE]
feat1   <- feat1[,store_total_quantity:=NULL]
clust1  <- clustering( title1, feat1 )

title2  <- "mkt share sell-in with total sell-in (no store dummies)"
feat2   <- stores_dummies_mktshares[,1:242, with=FALSE]
clust2  <- clustering( title2, feat2 )

title3  <- "mkt share sell-in with store dummies (no total sell-in)"
feat3   <- stores_dummies_mktshares[,1:1111, with=FALSE]
feat3   <- feat3[,store_total_quantity:=NULL]
clust3  <- clustering( title3, feat3 )

title4  <- "mkt share sell-in with total sell-in and store dummies"
feat4   <- stores_dummies_mktshares[,1:1111, with=FALSE]
clust4  <- clustering( title4, feat4 )

# all mkt share by shelf space variables
title5  <- "mkt share shelfspace (no total ss or store dummies)"
feat5   <- stores_dummies_mktshares[,1112:1353, with=FALSE]
feat5   <- feat5[,total_shelf_space:=NULL]
clust5  <- clustering( title5, feat5 )

title6  <- "mkt share shelfspace with total ss (no store dummies)"
feat6   <- stores_dummies_mktshares[,1112:1353, with=FALSE]
clust6  <- clustering( title6, feat6 )

title7  <- "mkt share shelfspace with store dummies (no total ss)"
feat7   <- stores_dummies_mktshares[,243:1353, with=FALSE]
feat7   <- feat7[,total_shelf_space:=NULL]
clust7  <- clustering( title7, feat7 )

title8  <- "mkt share shelfspace with total ss and store dummies"
feat8   <- stores_dummies_mktshares[,243:1353, with=FALSE]
clust8  <- clustering( title8, feat8 )

# all mkt share by sell-in AND all mkt share by shelf space variables - no store dummies
title9  <- "mkt share sell-in and shelfspace (no total sell-in, total ss or store dummies)"
feat9   <- stores_dummies_mktshares[,c(1:242, 1112:1353), with=FALSE]
feat9   <- feat9[,c("store_total_quantity", "total_shelf_space"):=NULL]
clust9  <- clustering( title9, feat9 )

title10  <- "mkt share sell-in and shelfspace with total sell-in (no total ss or store dummies)"
feat10   <- stores_dummies_mktshares[,c(1:242, 1112:1353), with=FALSE]
feat10   <- feat10[,total_shelf_space:=NULL]
clust10  <- clustering( title10, feat10 )

title11  <- "mkt share sell-in and shelfspace with total ss (no total sell-in or store dummies)"
feat11   <- stores_dummies_mktshares[,c(1:242, 1112:1353), with=FALSE]
feat11   <- feat11[,store_total_quantity:=NULL]
clust11  <- clustering( title11, feat11 )

title12  <- "mkt share sell-in and shelfspace with total sell-in and total ss (no store dummies)"
feat12   <- stores_dummies_mktshares[,c(1:242, 1112:1353), with=FALSE]
clust12  <- clustering( title12, feat12 )

# all mkt share by sell-in AND all mkt share by shelf space variables - with store dummies
title13  <- "mkt share sell-in and shelfspace with store dummies (no total sell-in or total ss)"
feat13   <- as.data.table(as.data.frame(stores_dummies_mktshares))
feat13   <- feat13[,c("store_total_quantity", "total_shelf_space"):=NULL]
clust13  <- clustering( title13, feat13 )

title14  <- "mkt share sell-in and shelfspace with total sell-in and store dummies (no total ss)"
feat14   <- as.data.table(as.data.frame(stores_dummies_mktshares))
feat14   <- feat14[,total_shelf_space:=NULL]
clust14  <- clustering( title14, feat14 )

title15  <- "mkt share sell-in and shelfspace with total ss and store dummies (no total sell-in)"
feat15   <- as.data.table(as.data.frame(stores_dummies_mktshares))
feat15   <- feat15[,store_total_quantity:=NULL]
clust15  <- clustering( title15, feat15 )

title16  <- "mkt share sell-in and shelfspace with total sell-in, total ss and store dummies"
feat16   <- as.data.table(as.data.frame(stores_dummies_mktshares))
clust16  <- clustering( title16, feat16 )

################################################################################
   # Clusterings using ONLY PRODUCT mkt share variables #
################################################################################
# product mkt share by sell-in 
title17  <- "PROD mkt share sell-in (no total sell-in or store dummies)"
feat17   <- stores_dummies_mktshares[,1:108, with=FALSE]
feat17   <- feat17[,store_total_quantity:=NULL]
clust17  <- clustering( title17, feat17 )

title18  <- "PROD mkt share sell-in with total sell-in (no store dummies)"
feat18   <- stores_dummies_mktshares[,1:108, with=FALSE]
clust18  <- clustering( title18, feat18 )

title19  <- "PROD mkt share sell-in with store dummies (no total sell-in)"
feat19   <- stores_dummies_mktshares[,c(1:108,243:1111), with=FALSE]
feat19   <- feat19[,store_total_quantity:=NULL]
clust19  <- clustering( title19, feat19 )

title20  <- "PROD mkt share sell-in with total sell-in and store dummies"
feat20   <- stores_dummies_mktshares[,c(1:108,243:1111), with=FALSE]
clust20  <- clustering( title20, feat20 )

# product mkt share by shelf space 
title21  <- "PROD mkt share shelfspace (no total ss or store dummies)"
feat21   <- stores_dummies_mktshares[,1112:1219, with=FALSE]
feat21   <- feat21[,total_shelf_space:=NULL]
clust21  <- clustering( title21, feat21 )

title22  <- "PROD mkt share shelfspace with total ss (no store dummies)"
feat22   <- stores_dummies_mktshares[,1112:1219, with=FALSE]
clust22  <- clustering( title22, feat22 )

title23  <- "PROD mkt share shelfspace with store dummies (no total ss)"
feat23   <- stores_dummies_mktshares[,243:1219, with=FALSE]
feat23   <- feat23[,total_shelf_space:=NULL]
clust23  <- clustering( title23, feat23 )

title24  <- "PROD mkt share shelfspace with total ss and store dummies"
feat24   <- stores_dummies_mktshares[,243:1219, with=FALSE]
clust24  <- clustering( title24, feat24 )

# product mkt share by sell-in AND product mkt share by shelf space - no store dummies
title25  <- "PROD mkt share sell-in and shelfspace (no total sell-in, total ss or store dummies)"
feat25   <- stores_dummies_mktshares[,c(1:108, 1112:1219), with=FALSE]
feat25   <- feat25[,c("store_total_quantity", "total_shelf_space"):=NULL]
clust25  <- clustering( title25, feat25 )

title26  <- "PROD mkt share sell-in and shelfspace with total sell-in (no total ss or store dummies)"
feat26   <- stores_dummies_mktshares[,c(1:108, 1112:1219), with=FALSE]
feat26   <- feat26[,total_shelf_space:=NULL]
clust26  <- clustering( title26, feat26 )

title27  <- "PROD mkt share sell-in and shelfspace with total ss (no total sell-in or store dummies)"
feat27   <- stores_dummies_mktshares[,c(1:108, 1112:1219), with=FALSE]
feat27   <- feat27[,store_total_quantity:=NULL]
clust27  <- clustering( title27, feat27 )

title28  <- "PROD mkt share sell-in and shelfspace with total sell-in and total ss (no store dummies)"
feat28   <- stores_dummies_mktshares[,c(1:108, 1112:1219), with=FALSE]
clust28  <- clustering( title28, feat28 )

# product mkt share by sell-in AND product mkt share by shelf space - with store dummies
title29  <- "PROD mkt share sell-in and shelfspace with store dummies (no total sell-in or total ss)"
feat29   <- stores_dummies_mktshares[,c(1:108, 243:1111, 1112:1219),with=FALSE]
feat29   <- feat29[,c("store_total_quantity", "total_shelf_space"):=NULL]
clust29  <- clustering( title29, feat29 )

title30  <- "PROD mkt share sell-in and shelfspace with total sell-in and store dummies (no total ss)"
feat30   <- stores_dummies_mktshares[,c(1:108, 243:1111, 1112:1219),with=FALSE]
feat30   <- feat30[,total_shelf_space:=NULL]
clust30  <- clustering( title30, feat30 )

title31  <- "PROD mkt share sell-in and shelfspace with total ss and store dummies (no total sell-in)"
feat31   <- stores_dummies_mktshares[,c(1:108, 243:1111, 1112:1219),with=FALSE]
feat31   <- feat31[,store_total_quantity:=NULL]
clust31  <- clustering( title31, feat31 )

title32  <- "PROD mkt share sell-in and shelfspace with total sell-in, total ss and store dummies"
feat32   <- stores_dummies_mktshares[,c(1:108, 243:1111, 1112:1219),with=FALSE]
clust32  <- clustering( title32, feat32 )

################################################################################
   # Clusterings using ONLY PRODUCTVAR mkt share variables #
################################################################################
# prodvar mkt share variables by sell-in 
title33  <- "PRODVAR mkt share sell-in (no total sell-in or store dummies)"
feat33   <- stores_dummies_mktshares[,108:242, with=FALSE]
feat33   <- feat33[,store_total_quantity:=NULL]
clust33  <- clustering( title33, feat33 )

title34  <- "PRODVAR mkt share sell-in with total sell-in (no store dummies)"
feat34   <- stores_dummies_mktshares[,108:242, with=FALSE]
clust34  <- clustering( title34, feat34 )

title35  <- "PRODVAR mkt share sell-in with store dummies (no total sell-in)"
feat35   <- stores_dummies_mktshares[,c(108:242,243:1111), with=FALSE]
feat35   <- feat35[,store_total_quantity:=NULL]
clust35  <- clustering( title35, feat35 )

title36  <- "PRODVAR mkt share sell-in with total sell-in and store dummies"
feat36   <- stores_dummies_mktshares[,c(108:242,243:1111), with=FALSE]
clust36  <- clustering( title36, feat36 )

# prodvar mkt share variables by shelf space 
title37  <- "PRODVAR mkt share shelfspace (no total ss or store dummies)"
feat37   <- stores_dummies_mktshares[,c(1112, 1220:1353), with=FALSE]
feat37   <- feat37[,total_shelf_space:=NULL]
clust37  <- clustering( title37, feat37 )

title38  <- "PRODVAR mkt share shelfspace with total ss (no store dummies)"
feat38   <- stores_dummies_mktshares[,c(1112, 1220:1353), with=FALSE]
clust38  <- clustering( title38, feat38 )

title39  <- "PRODVAR mkt share shelfspace with store dummies (no total ss)"
feat39   <- stores_dummies_mktshares[,c(1112, 243:1111, 1220:1353), with=FALSE]
feat39   <- feat39[,total_shelf_space:=NULL]
clust39  <- clustering( title39, feat39 )

title40  <- "PRODVAR mkt share shelfspace with total ss and store dummies"
feat40   <- stores_dummies_mktshares[,c(1112, 243:1111, 1220:1353), with=FALSE]
clust40  <- clustering( title40, feat40 )

# prodvar mkt share by sell-in variables AND prodvar mkt share by shelf space variables - no store dummies
title41  <- "PRODVAR mkt share sell-in and shelfspace (no total sell-in, total ss or store dummies)"
feat41   <- stores_dummies_mktshares[,c(108:242, 1112, 1220:1353), with=FALSE]
feat41   <- feat41[,c("store_total_quantity", "total_shelf_space"):=NULL]
clust41  <- clustering( title41, feat41 )

title42  <- "PRODVAR mkt share sell-in and shelfspace with total sell-in (no total ss or store dummies)"
feat42   <- stores_dummies_mktshares[,c(108:242, 1112, 1220:1353), with=FALSE]
feat42   <- feat42[,total_shelf_space:=NULL]
clust42  <- clustering( title42, feat42 )

title43  <- "PRODVAR mkt share sell-in and shelfspace with total ss (no total sell-in or store dummies)"
feat43   <- stores_dummies_mktshares[,c(108:242, 1112, 1220:1353), with=FALSE]
feat43   <- feat43[,store_total_quantity:=NULL]
clust43  <- clustering( title43, feat43 )

title44  <- "PRODVAR mkt share sell-in and shelfspace with total sell-in and total ss (no store dummies)"
feat44   <- stores_dummies_mktshares[,c(108:242, 1112, 1220:1353), with=FALSE]
clust44  <- clustering( title44, feat44 )

# prodvar mkt share by sell-in variables AND prodvar mkt share by shelf space variables  - with store dummies
title45  <- "PRODVAR mkt share sell-in and shelfspace with store dummies (no total sell-in or total ss)"
feat45   <- stores_dummies_mktshares[,c(108:242, 243:1111, 1112, 1220:1353),with=FALSE]
feat45   <- feat45[,c("store_total_quantity", "total_shelf_space"):=NULL]
clust45  <- clustering( title45, feat45 )

title46  <- "PRODVAR mkt share sell-in and shelfspace with total sell-in and store dummies (no total ss)"
feat46   <- stores_dummies_mktshares[,c(108:242, 243:1111, 1112, 1220:1353),with=FALSE]
feat46   <- feat46[,total_shelf_space:=NULL]
clust46  <- clustering( title46, feat46 )

title47  <- "PRODVAR mkt share sell-in and shelfspace with total ss and store dummies (no total sell-in)"
feat47   <- stores_dummies_mktshares[,c(108:242, 243:1111, 1112, 1220:1353),with=FALSE]
feat47   <- feat47[,store_total_quantity:=NULL]
clust47  <- clustering( title47, feat47 )

title48  <- "PRODVAR mkt share sell-in and shelfspace with total sell-in, total ss and store dummies"
feat48   <- stores_dummies_mktshares[,c(108:242, 243:1111, 1112, 1220:1353),with=FALSE]
clust48  <- clustering( title48, feat48 )


```

## All all market share variables (product, subFam, fam, etc)
#### All mkt share by sell-in variables and variants
```{r, echo=FALSE}
# plot different clusterings
grid.arrange(clust1$plot, clust2$plot, clust3$plot, clust4$plot, ncol = 2)
```


#### All mkt share by shelf space variables and variants
```{r, echo=FALSE}
grid.arrange(clust5$plot, clust6$plot, clust7$plot, clust8$plot, ncol = 2)
```


#### All mkt share by sell-in and shelf space variables and variants (without store dummies)
```{r, echo=FALSE}
grid.arrange(clust9$plot, clust10$plot, clust11$plot, clust12$plot, ncol = 2)
```


#### All mkt share by sell-in and shelf space variables and variants (with store dummies)
```{r, echo=FALSE}
grid.arrange(clust13$plot, clust14$plot, clust15$plot, clust16$plot, ncol = 2)
```


## Only PRODUCT market share variables 
#### PRODUCT mkt share by sell-in and variants
```{r, echo=FALSE}
# plot different clusterings
grid.arrange(clust17$plot, clust18$plot, clust19$plot, clust20$plot, ncol = 2)
```


#### PRODUCT mkt share by shelf space and variants
```{r, echo=FALSE}
grid.arrange(clust21$plot, clust22$plot, clust23$plot, clust24$plot, ncol = 2)
```


#### PRODUCT mkt share by sell-in and shelf space and variants (without store dummies)
```{r, echo=FALSE}
grid.arrange(clust25$plot, clust26$plot, clust27$plot, clust28$plot, ncol = 2)
```


#### PRODUCT mkt share by sell-in and shelf space and variants (with store dummies)
```{r, echo=FALSE}
grid.arrange(clust29$plot, clust30$plot, clust31$plot, clust32$plot, ncol = 2)
```


## Only PRODVAR market share variables 
#### PRODVAR mkt share by sell-in and variants
```{r, echo=FALSE}
# plot different clusterings
grid.arrange(clust33$plot, clust34$plot, clust35$plot, clust36$plot, ncol = 2)
```


#### PRODVAR mkt share by shelf space and variants
```{r, echo=FALSE}
grid.arrange(clust37$plot, clust38$plot, clust39$plot, clust40$plot, ncol = 2)
```


#### PRODVAR mkt share by sell-in and shelf space and variants (without store dummies)
```{r, echo=FALSE}
grid.arrange(clust41$plot, clust42$plot, clust43$plot, clust44$plot, ncol = 2)
```


#### PRODVAR mkt share by sell-in and shelf space and variants (with store dummies)
```{r, echo=FALSE}
grid.arrange(clust45$plot, clust46$plot, clust47$plot, clust48$plot, ncol = 2)
```